Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-12-05 manuscripts/manuscript.Rmd
edit readmes and tidy up

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (beginning with
@Clark1976, @Clark2009 and @Walters1978)
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of
understanding how to manage human intervention into natural systems.
Simultaneously, there has been increasing recognition of the importance of
multiple steady states or `tipping points' [@Scheffer2001; @Scheffer2009;
@Polasky2011] in ecological systems.

@@

### Parametric models

@@

\subsubsection{Parametric models}\label{parametric-models}

@@

For the prior on  $\ell$,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.  The prior  on $\sigma$,  $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.

>>>>

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (beginning with
@Clark1976, @Clark2009 and @Walters1978)
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of
understanding how to manage human intervention into natural systems.
Simultaneously, there has been increasing recognition of the importance of
multiple steady states or 'tipping points' [@Scheffer2001; @Scheffer2009;
@Polasky2011] in ecological systems.

@@

@@

### Parametric Models ###

@@

For the prior on  $\ell$,
$\alpha =$ `r d.p[1]` and 
$\beta =$ `r d.p[2]`. 
The prior  on $\sigma$,
$\alpha =$ `r s2.p[1]`
and $\beta =$ `r s2.p[2]`.

### 2014-12-05 manuscripts/manuscript.tex
edit readmes and tidy up

\subsubsection{Parametric models}\label{parametric-models}

We let $X(t)$ denote the size (numbers or biomass) of the focal
population at time $t$ and assume that in the absence of take its
dynamics are:

@@

Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
and $\mathbf{p}$ is a vector of parameters to be estimated from the
data. We describe the three choices for $f(X(t),\mathbf{p})$ in the next
section.\\The control action is a harvest or take, $h(t)$, measured in
the same units as $X$, at time $t$. Thus, in the presence of take, the
population size on the right hand side of Eqn 1 is replaced by
$S(t) = X(t) - h(t)$.

To construct the value function, we consider a return when $X(t) = x(t)$
and harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example,
if the return is the harvest at time $t$, then
$R(x(t), h(t)) = \min(x(t), h(t))$. We assume that future harvests are
discounted relative to current ones at a constant rate of discount
$\delta$ and ask for the harvest policy that maximizes total discounted
harvest between the current time $t$ and a final time $T$. That is, we
seek to maximize over choices of harvest
$E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the state
dynamics are given by Eqn 1 and $E$ denotes the expectation over future
population states.

In order to find that policy, we introduce the value function
$V(x(t), t)$ representing the total discounted catch from time $t$
onwards given that $X(t) = x(t)$. This value function satisfies an
equation of SDP (Mangel and Clark 1988, Clark and Mangel 2000, Clark
2009 Mangel 2014),

@@

where expectation is taken over all possible values of the next state,
$X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
is applied, the immediate return is $R(h(t), x(t))$. When the sole
source of uncertainty is the process stochasticity term, $Z$, and thus
the expectation in Eqn 2 is equivalent to taking expectations over
$Z(t)$. That is

@@

where the population size after the take is $x(t) - h(t)$, which is then
translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by
$Z(t) f(x(t)-h(t)|\mathbf{p})$).

@@

When the underlying population dynamics are unknown (the case of
structural uncertainty), the function $f$ itself is uncertain and the
expectation for the next state includes uncertainty in $f$ as well. That
is

@@

We consider the finite time problem with $T=$ 1000, which we solve using
the standard value iteration algorithm (see Mangel and Clark 1988, Clark
and Mangel 2000).

\subsubsection{Parametric models}\label{parametric-models}

We consider three candidate parametric models for the population
dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), and
the Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers}.
In all three, we let $K$ denote the carrying capacity and $r$ the
maximum per capita growth rate. The Ricker model has two parameters and
the right hand side of Eqn 1 is

@@

where $X_c$ denotes the location of the unstable steady state (i.e., the
tipping point).

@@

where $\theta = 1$ corresponds to Beverton-Holt dynamics and $\theta$
\textgreater{} 2 leads to Allee effects and multiple stable states.

The Ricker model does not lead to multiple steady states. The Allen
model resembles the Ricker dynamics with an added Allee effect parameter
(Courchamp et al. 2008), below which the population cannot persist. The
Myers model also has three parameters and contains an Allee threshold,
but unlike the Ricker model saturates at high population size. The
multiplicative log-normal stochasticity in Eqn 1 introduces one
additional parameter $\sigma$ that must be estimated.

@@

A GP is a stochastic process for which any realization consisting of n
points follows a multivariate normal distribution of dimension $n$. To
characterize the GP we need a mean function and a covariance function.
We proceed as follows.

As before, we assume that the data $X(t)$ are observed with process
stochasticity around a mean function $g(X(t))$

@@

where $\varepsilon$ are IID normal random variables with zero-mean and
variance $\sigma^2$. Note that we have chosen to assume additive
stochasticity. While we could consider log-normal stochasticity as in
the parametric models, we make this choice to emphasize that the
Gaussian process approach need not have structurally correct
stochasticity to be effective.

In order to make predictions, we update the GP based on the observed set
of transitions. To do so, we collect the time series of observed states
into a vector of ``current'' states,
$\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}$ and a vector of
``next'' states $\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}$ where
$T$ is the time of the final observation. Conditional on these
observations, the predicted next state, $g(X_p)$, for any given
``current'' state, $X_p$ follows a normal distribution with mean $E$ and
variance $C$ determined using the standard rules for conditioning in
multivariate normals, i.e.

@@

Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
with ones down the diagonal and zeros elsewhere) and $K$ is the
`covariance kernel.' The covariance kernel controls how much influence
one observation has on another. In the present application we use the
squared-exponential kernel which, when evaluated over a pair of vectors,
say $\mathbf{x}$ and $\mathbf{y}$, generates a covariance matrix whose
$i,j$th element is given by

@@

so that $\ell$ gives the characteristic length-scale over which
correlation between two observations decays. See Rasmussen and Williams
(2006) for other choices of covariance kernels and their properties.
Note that this simple formulation assumes a prior mean of zero. For the
parameters we use inverse Gamma priors on both the length-scale $\ell$
and $\sigma$, thus for example

@@

For the prior on $\ell$, $\alpha = $ 10 and $\beta = $ 10. The prior on
$\sigma$, $\alpha = $ 5 and $\beta = $ 5.

@@

We derive the harvest policy from the estimated GP by inserting it into
a SDP algorithm. Given the GP posteriors, we construct the transition
matrix representing the probability of going to each state $X(t+1)$
given any current state $x(t)$ and any harvest $h(t)$ (See the function
\texttt{gp\_transition\_matrix()} in the provided R package). Given this
transition matrix, we use the same value iteration algorithm as in the
parametric case to determine the optimal policy. In doing so, the
uncertainty in the future state under the GP, $X(t+1)$, includes both
process uncertainty (based on the estimation of $\sigma$) and structural
uncertainty of the posterior collection of curves.

@@

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.12\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\centering
Allen
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Ricker
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Myers
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{DIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
50.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.45
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.41
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{AIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-24.51
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-30.13
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-27.01
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{BIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-17.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-25.06
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-20.25
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Model selection scores for several common criteria (DIC:
Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
Bayesian Information Criterion) all select the wrong model. As the true
(Allen) model is not distinguishable from the simpler (Ricker) model in
the region of the observed data, this error cannot be avoided regardless
of the model choice criterion. This highlights the danger of model
choice when the selected model will be used outside of the observed
range of the data.}
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function $f$ under each model. Plots show replicates drawn from
the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3-1.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S(t)$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979) Several policies show a numerical
jitter due to the discretization of states in the dynamic programming
algorithm; doubling the number of grid points did not qualitatively
change the results.}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely (Figure 3). Policies
are shown in terms of target escapement, $S(t) = x_t - h$. Under models
such as this a constant escapement policy is expected to be optimal
(Reed 1979), whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return
the population to $S$. Whenever a model predicts that the population
will not persist below a certain threshold, the optimal solution is to
harvest the entire population immediately, resulting in an escapement
$S=0$, as seen in the true (correct form, exact parameters) model, the
Allen model (correct form, estimated parameters) and the GP. Only the
structurally correct model (Allen model) and the GP produce policies
close to the true optimum policy.

@@

The GPDP is only weakly influenced by increasing stochasticity or
increasing Allee effects over much of the range (Figure S2). Larger
$\sigma$ or higher Allee levels make even the optimal solution without
any model or parameter uncertainty unable to harvest the population
effectively (e.g.~the stochasticity is large enough to violate the
self-sustaining criterion of Reed (1979)).

@@

Our results illustrate that model-choice approaches can be absolutely
misleading -- by providing support to models that cannot capture tipping
point dynamics because they have fewer parameters and the data are far
from the tipping point. That is, when the data come from around the
stable steady state, all the parametric models are approximately linear
and approximately identical. Thus, it is intuitive that all model
selection methods choose the simplest model. In a complex world, the
result is suboptimal. But in a world that might contain tipping points,
the result could be disastrous.

@@

Outside of the observed range of the data, the GP reverts to the prior,
and consequently the choice of the prior can also play a significant
role in determining the optimal policy. In the examples shown here we
have selected a prior that is both relatively uninformative (due to the
broad priors placed on its parameters $\ell$ and $\sigma$ and simple
(the choice of our covariance function, Eqns 12 and 13 ). In practice,
these should be chosen to confer particular biological properties. In
principle, this may allow a manager to improve the performance of the
GPDP by adding detail as is justified. For instance, it would be
possible to use a linear or a Ricker-shaped mean in the prior without
making the much stronger assumption that the Ricker is the structurally
correct model (Sugeno and Munch 2013a). One fruitful avenue of future
research is identifying criteria to ensure the prior and the reward
function are chosen appropriately for the problem at hand.

>>>>

We let \(X(t)\) denote the size (numbers or biomass) of the focal
population at time \(t\) and assume that in the absence of take its
dynamics are:

@@

Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
and \(\mathbf{p}\) is a vector of parameters to be estimated from the
data. We describe the three choices for \(f(X(t),\mathbf{p})\) in the
next section.\\The control action is a harvest or take, \(h(t)\),
measured in the same units as \(X\), at time \(t\). Thus, in the
presence of take, the population size on the right hand side of Eqn 1 is
replaced by \(S(t) = X(t) - h(t)\).

To construct the value function, we consider a return when
\(X(t) = x(t)\) and harvest \(h(t) = h\) denoted as the reward,
\(R(x(t), h)\). For example, if the return is the harvest at time \(t\),
then \(R(x(t), h(t)) = \min(x(t), h(t))\). We assume that future
harvests are discounted relative to current ones at a constant rate of
discount \(\delta\) and ask for the harvest policy that maximizes total
discounted harvest between the current time \(t\) and a final time
\(T\). That is, we seek to maximize over choices of harvest
\(E [ \sum_{t = 0}^{T} R(X(t), h(t), t) \delta^t]\), where the state
dynamics are given by Eqn 1 and \(E\) denotes the expectation over
future population states.

In order to find that policy, we introduce the value function
\(V(x(t), t)\) representing the total discounted catch from time \(t\)
onwards given that \(X(t) = x(t)\). This value function satisfies an
equation of SDP (Mangel and Clark 1988, Clark and Mangel 2000, Clark
2009 , Mangel 2014),

@@

where expectation is taken over all possible values of the next state,
\(X(t+1)\), and maximized over all possible choices of harvest,
\(h(t)\). That is, at time \(t\), when population size is \(x(t)\) and
harvest \(h(t)\) is applied, the immediate return is \(R(h(t), x(t))\).
When the sole source of uncertainty is the process stochasticity term,
\(Z\), and thus the expectation in Eqn 2 is equivalent to taking
expectations over \(Z(t)\). That is

@@

where the population size after the take is \(x(t) - h(t)\), which is
then translated into \(X(t+1)\) by Eqn 1 (that is, we replace \(X(t+1)\)
by \(Z(t) f(x(t)-h(t)|\mathbf{p})\)).

@@

When the underlying population dynamics are unknown (the case of
structural uncertainty), the function \(f\) itself is uncertain and the
expectation for the next state includes uncertainty in \(f\) as well.
That is

@@

We consider the finite time problem with \(T=\) 1000, which we solve
using the standard value iteration algorithm (see Mangel and Clark 1988,
Clark and Mangel 2000).

\subsubsection{Parametric Models}\label{parametric-models}

We consider three candidate parametric models for the population
dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), and
the Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers}.
In all three, we let \(K\) denote the carrying capacity and \(r\) the
maximum per capita growth rate. The Ricker model has two parameters and
the right hand side of Eqn 1 is

@@

where \(X_c\) denotes the location of the unstable steady state (i.e.,
the tipping point).

@@

where \(\theta = 1\) corresponds to Beverton-Holt dynamics and
\(\theta\) \textgreater{} 2 leads to Allee effects and multiple stable
states.

The Ricker model does not lead to multiple steady states. The Allen
model resembles the Ricker dynamics with an added Allee effect parameter
(Courchamp et al. 2008), below which the population cannot persist. The
Myers model also has three parameters and contains an Allee threshold,
but unlike the Ricker model saturates at high population size. The
multiplicative log-normal stochasticity in Eqn 1 introduces one
additional parameter \(\sigma\) that must be estimated.

@@

A GP is a stochastic process for which any realization consisting of n
points follows a multivariate normal distribution of dimension \(n\). To
characterize the GP we need a mean function and a covariance function.
We proceed as follows.

As before, we assume that the data \(X(t)\) are observed with process
stochasticity around a mean function \(g(X(t))\)

@@

where \(\varepsilon\) are IID normal random variables with zero-mean and
variance \(\sigma^2\). Note that we have chosen to assume additive
stochasticity. While we could consider log-normal stochasticity as in
the parametric models, we make this choice to emphasize that the
Gaussian process approach need not have structurally correct
stochasticity to be effective.

In order to make predictions, we update the GP based on the observed set
of transitions. To do so, we collect the time series of observed states
into a vector of ``current'' states,
\(\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}\) and a vector of
``next'' states \(\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}\)
where \(T\) is the time of the final observation. Conditional on these
observations, the predicted next state, \(g(X_p)\), for any given
``current'' state, \(X_p\) follows a normal distribution with mean \(E\)
and variance \(C\) determined using the standard rules for conditioning
in multivariate normals, i.e.

@@

NA

@@

so that \(\ell\) gives the characteristic length-scale over which
correlation between two observations decays. See Rasmussen and Williams
(2006) for other choices of covariance kernels and their properties.
Note that this simple formulation assumes a prior mean of zero. For the
parameters we use inverse Gamma priors on both the length-scale \(\ell\)
and \(\sigma\), thus for example

@@

For the prior on \(\ell\), \(\alpha =\) 10 and \(\beta =\) 10. The prior
on \(\sigma\), \(\alpha =\) 5 and \(\beta =\) 5.

@@

We derive the harvest policy from the estimated GP by inserting it into
a SDP algorithm. Given the GP posteriors, we construct the transition
matrix representing the probability of going to each state \(X(t+1)\)
given any current state \(x(t)\) and any harvest \(h(t)\) (See the
function \texttt{gp\_transition\_matrix()} in the provided R package).
Given this transition matrix, we use the same value iteration algorithm
as in the parametric case to determine the optimal policy. In doing so,
the uncertainty in the future state under the GP, \(X(t+1)\), includes
both process uncertainty (based on the estimation of \(\sigma\)) and
structural uncertainty of the posterior collection of curves.

@@

\begin{longtable}[c]{@{}cccc@{}}
\caption{Model selection scores for several common criteria (DIC:
Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
Bayesian Information Criterion) all select the wrong model. As the true
(Allen) model is not distinguishable from the simpler (Ricker) model in
the region of the observed data, this error cannot be avoided regardless
of the model choice criterion. This highlights the danger of model
choice when the selected model will be used outside of the observed
range of the data.}\\
\toprule\addlinespace
\begin{minipage}[b]{0.12\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\centering
Allen
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Ricker
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Myers
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{DIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
50.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.45
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.41
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{AIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-24.51
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-30.13
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-27.01
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{BIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-17.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-25.06
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-20.25
\end{minipage}
\\\addlinespace
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function \(f\) under each model. Plots show replicates drawn
from the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3-1.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
\(S(t)\), as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979) Several policies show a numerical
jitter due to the discretization of states in the dynamic programming
algorithm; doubling the number of grid points did not qualitatively
change the results.}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely (Figure 3). Policies
are shown in terms of target escapement, \(S(t) = x_t - h\). Under
models such as this a constant escapement policy is expected to be
optimal (Reed 1979), whereby population levels below a certain size
\(S\) are unharvested, while above that size the harvest strategy aims
to return the population to \(S\). Whenever a model predicts that the
population will not persist below a certain threshold, the optimal
solution is to harvest the entire population immediately, resulting in
an escapement \(S=0\), as seen in the true (correct form, exact
parameters) model, the Allen model (correct form, estimated parameters)
and the GP. Only the structurally correct model (Allen model) and the GP
produce policies close to the true optimum policy.

@@

The GPDP is only weakly influenced by increasing stochasticity or
increasing Allee effects over much of the range (Figure S2). Larger
\(\sigma\) or higher Allee levels make even the optimal solution without
any model or parameter uncertainty unable to harvest the population
effectively (e.g.~the stochasticity is large enough to violate the
self-sustaining criterion of Reed (1979)).

@@

Our results illustrate that model-choice approaches can be absolutely NA
dynamics because they have fewer parameters and the data are far from
the tipping point. That is, when the data come from around the stable
steady state, all the parametric models are approximately linear and
approximately identical. Thus, it is intuitive that all model selection
methods choose the simplest model. In a complex world, the result is
suboptimal. But in a world that might contain tipping points, the result
could be disastrous.

@@

Outside of the observed range of the data, the GP reverts to the prior,
and consequently the choice of the prior can also play a significant
role in determining the optimal policy. In the examples shown here we
have selected a prior that is both relatively uninformative (due to the
broad priors placed on its parameters \(\ell\) and \(\sigma\) and simple
(the choice of our covariance function, Eqns 12 and 13 ). In practice,
these should be chosen to confer particular biological properties. In
principle, this may allow a manager to improve the performance of the
GPDP by adding detail as is justified. For instance, it would be
possible to use a linear or a Ricker-shaped mean in the prior without
making the much stronger assumption that the Ricker is the structurally
correct model (Sugeno and Munch 2013a). One fruitful avenue of future
research is identifying criteria to ensure the prior and the reward
function are chosen appropriately for the problem at hand.

### 2014-12-05 manuscripts/submit.tex
tweak tex paths for submission engine
edit readmes and tidy up

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.12\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\centering
Allen
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Ricker
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Myers
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{DIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
50.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.45
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.41
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{AIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-24.51
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-30.13
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-27.01
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{BIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-17.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-25.06
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-20.25
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Model selection scores for several common criteria (DIC:
Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
Bayesian Information Criterion) all select the wrong model. As the true
(Allen) model is not distinguishable from the simpler (Ricker) model in
the region of the observed data, this error cannot be avoided regardless
of the model choice criterion. This highlights the danger of model
choice when the selected model will be used outside of the observed
range of the data.}
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function $f$ under each model. Plots show replicates drawn from
the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3-1.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S(t)$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979) Several policies show a numerical
jitter due to the discretization of states in the dynamic programming
algorithm; doubling the number of grid points did not qualitatively
change the results.}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely (Figure 3). Policies
are shown in terms of target escapement, $S(t) = x_t - h$. Under models
such as this a constant escapement policy is expected to be optimal
(Reed 1979), whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return
the population to $S$. Whenever a model predicts that the population
will not persist below a certain threshold, the optimal solution is to
harvest the entire population immediately, resulting in an escapement
$S=0$, as seen in the true (correct form, exact parameters) model, the
Allen model (correct form, estimated parameters) and the GP. Only the
structurally correct model (Allen model) and the GP produce policies
close to the true optimum policy.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4-1.jpeg}
\caption{In the management context, GPDP outperforms approaches based on
parametric models. We show 100 replicate simulations of the stock
dynamics (Eqn 1) under the policies derived from each of the estimated
models, as well as the policy based on the exact underlying model.}
\end{figure}

>>>>

\begin{longtable}[c]{@{}cccc@{}}
\caption{Model selection scores for several common criteria (DIC:
Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
Bayesian Information Criterion) all select the wrong model. As the true
(Allen) model is not distinguishable from the simpler (Ricker) model in
the region of the observed data, this error cannot be avoided regardless
of the model choice criterion. This highlights the danger of model
choice when the selected model will be used outside of the observed
range of the data.}\\
\toprule\addlinespace
\begin{minipage}[b]{0.12\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\centering
Allen
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Ricker
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
Myers
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{DIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
50.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.45
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
50.41
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{AIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-24.51
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-30.13
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-27.01
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.12\columnwidth}\centering
\textbf{BIC}
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
-17.75
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-25.06
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
-20.25
\end{minipage}
\\\addlinespace
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript-figure_2-1.jpeg}
\caption{The inferred Gaussian process compared to the true process and
maximum-likelihood estimated (MLE) process. We show the expected value
for the function \(f\) under each model. Plots show replicates drawn
from the posterior distributions in order to convey uncertainty of the
estimates. Note the MLE is a point estimate of parameters and so
reflects no uncertainty in the distribution. The training data are also
shown as black points. The GP is conditioned on (0,0), shown as a
pseudo-data point.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3-1.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
\(S(t)\), as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979) Several policies show a numerical
jitter due to the discretization of states in the dynamic programming
algorithm; doubling the number of grid points did not qualitatively
change the results.}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely (Figure 3). Policies
are shown in terms of target escapement, \(S(t) = x_t - h\). Under
models such as this a constant escapement policy is expected to be
optimal (Reed 1979), whereby population levels below a certain size
\(S\) are unharvested, while above that size the harvest strategy aims
to return the population to \(S\). Whenever a model predicts that the
population will not persist below a certain threshold, the optimal
solution is to harvest the entire population immediately, resulting in
an escapement \(S=0\), as seen in the true (correct form, exact
parameters) model, the Allen model (correct form, estimated parameters)
and the GP. Only the structurally correct model (Allen model) and the GP
produce policies close to the true optimum policy.

\begin{figure}[htbp]
\centering
\includegraphics{manuscript-figure_4-1.jpeg}
\caption{In the management context, GPDP outperforms approaches based on
parametric models. We show 100 replicate simulations of the stock
dynamics (Eqn 1) under the policies derived from each of the estimated
models, as well as the policy based on the exact underlying model.}
\end{figure}
